Leader functions and qualities across 59 cultures: an exploratory analysis

### 2019-11-28 leadership_across_cultures_contexts.Rmd
Edits to MS.

Figure \@ref(fig:qualcluster) displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:funcluster) displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables. Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data. We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $>95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).

```{r qualcluster, fig.cap="Cluster analysis of leader quality variables. Distances are $1 - cor$. Ward agglomeration method. p-values computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
# plot(m_pvclust_qual)
# pvrect(m_pvclust_qual, alpha = 0.9)
knitr::include_graphics("Figures/m_pvclust_qual.final.png")
```

The leader quality cluster analysis (Figure \@ref(fig:qualcluster)) idenfitified two top-level clusters. Within each top-level cluster we identify two interpretable and generally strongly supported sub-clusters which are annotated with text. Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (e.g., *Authority*, *Aggressiveness*, *Feared*) and shamanism (e.g., *Supernatural*, *Artistic performance*). The second sub-cluster includes qualities of leaders related to material, social, and reproductive success.

```{r funcluster, fig.cap="Cluster analysis of leader function variables. Distances are $1 - cor$. Ward agglomeration method. p-values computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
# plot(m_pvclust_fun)
# pvrect(m_pvclust_fun, alpha = 0.9)
knitr::include_graphics("Figures/m_pvclust_fun.final.png")
```

@@

A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance. We draw on the first two leader quality components for interpretation and additional analyses. We identify the first component (PC1) as a *Prestige vs. Dominance* measure. Low values on Leader qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*. High values are strongly associated with *Polygnous* and *Authority* and moderately associated with *High quality spouse*. Therefore, the "dominance" side of the *Prestige vs. Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:qualcluster)). Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component. See Figure \@ref(fig:qualitiesPCplot). The second leader qualities component we identify as an *In-group favoritism vs. Social Status* measure. Low values on the Leader qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*. High Values are strongly associated with *High status* and moderately associated with *Fairness* and *Age* and *Decisiveness/decision-making*.

>>>>

Figure \@ref(fig:cluster)A displays a dendrogram from the cluster analysis of the `r length(quality_vars)` leader quality variables and Figure \@ref(fig:cluster)B displays a dendrogram from the cluster analysis of the `r length(function_vars)` leader function variables. Each cluster analysis includes two estimates of significance for how strongly each cluster within the dendrogram is supported by the data. We rely on the AU (Approximately Unbiased) *p* values, which are computed by multiscale bootstrap resampling and represented as percentages (e.g., clusters with AU values $>95$ are strongly supported, and the top-level clusters are automatically highlighted by rectangles).

```{r cluster, fig.cap="Cluster analysis of leader quality variables (left) and leader function variables (right). Distances are $1 - cor$. Ward agglomeration method. p-values computed with 10,000 bootstrap samples using the pvclust package [@pvclust2015]."}
# plot(m_pvclust_qual)
# pvrect(m_pvclust_qual, alpha = 0.9)
knitr::include_graphics("Figures/m_pvclust.both.png")
```

The leader quality cluster analysis (Figure \@ref(fig:cluster)A) idenfitified two top-level clusters. Within each top-level cluster we identify two interpretable and generally strongly supported sub-clusters which are annotated with text. Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (e.g., *Authority*, *Aggressiveness*, *Feared*) and shamanism (e.g., *Supernatural*, *Artistic performance*). The second sub-cluster includes qualities of leaders related to material, social, and reproductive success.

@@

A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance. We draw on the first two leader quality components for interpretation and additional analyses. We identify the first component (PC1) as a *Prestige vs. Dominance* measure. Low values on Leader qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*. High values are strongly associated with *Polygnous* and *Authority* and moderately associated with *High quality spouse*. Therefore, the "dominance" side of the *Prestige vs. Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:cluster)). Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component. See Figure \@ref(fig:qualitiesPCplot). The second leader qualities component we identify as an *In-group favoritism vs. Social Status* measure. Low values on the Leader qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*. High Values are strongly associated with *High status* and moderately associated with *Fairness* and *Age* and *Decisiveness/decision-making*.
